---
title: "p8105_hw02_tab2187"
output: html_document
date: "2022-09-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(ccollapse = TRUE, message = FALSE)
```
##Problem 0
Created a git repository and figured out how to knit on save. This did not work for me on the last assignment so I am glad I was able to make it work on this one.

```{r}
library(tidyverse)
library(readxl)
library(readr)
options(tibble.print_min = 5)
```

##Problem 1
The process begins with data import, updates variable names, and selects the columns that will be used in later parts of this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
NYC_Subway=
  read_csv("NYC_Subway.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry =="Yes", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
NYC_Subway %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a data frame in which the number of rows is the number of ADA compliant stations. 

```{r}
NYC_Subway %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
NYC_Subway %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain data frames with the required stations in rows).

```{r}
NYC_Subway %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYC_Subway %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

##Problem 2
Specifying the sheet in excel that needs to be read is Mr. Trash Wheel. Omitted rows with notes, figures, and columns containing notes as well as columns that did not include dumpster specific data. Sports Balls data is now rounded using as.integer

```{r}
Mr_Trash_Wheel=
  readxl::read_excel("Trash_Wheel.xlsx", "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  select(-date)
```

Specifying the sheet in excel that needs to be read is Professor Trash Wheel. Omitted rows with notes, figures, and columns containing notes as well as columns that did not include dumpster specific data.

```{r}
Professor_Trash_Wheel=
  readxl::read_excel("Trash_Wheel.xlsx", "Professor Trash Wheel", range = "A2:M97") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  select(-date)
```

Merging the data sets Mr_Trash_Wheel and Professor_Trash_Wheel. When we first merge the data we cannot tell which data belongs to which data set. So we merge them with the id pathway. Once we merge them we see that 1 under the which_trash_wheel is Mr.Trash Wheel and 2 under the which_trash_wheel is Professor Trash Wheel. 

```{r}
Mr_Trash_Wheel$dumpster = as.double(Mr_Trash_Wheel$dumpster)
Mr_Trash_Wheel$year = as.double(Mr_Trash_Wheel$year)

TrashWheel_combined = 
  bind_rows(Mr_Trash_Wheel, Professor_Trash_Wheel, .id="which_trash_wheel")
```

##Inline R Code for Problem 2
The total weight of trash collected by Professor Trash Wheel= 
`r sum(Professor_Trash_Wheel$weight_tons)`

The total number of sports balls collected by Mr. Trash Wheel in 2020=
`r sum(Mr_Trash_Wheel$sports_balls)`

##Problem 3
First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r}
Poll=
  read_csv("pols-month.csv") %>% 
  janitor::clean_names() %>% 
   separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(prez_dem = recode(prez_dem, `1` = "dem", `0` = "gop",))
  mutate(prez_gop = recode(prez_gop, `1` = "gop", `0` = "dem", `2` = "unexpected"))  
pivot_longer(
  cols = start_with("prez"),
  names_to = NULL,
  values_to = "president") %>%
select(-day)
```
Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
SNP=
  read_csv("snp.csv") %>% 
  janitor::clean_names() %>% 
   separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate(month = month.name[month])
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
Unemployment = 
  read_csv("unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    january:december,
    names_to = "month", 
    values_to = "unemployment rates")
```

Join the datasets by merging snp into pols, and merging unemployment into the result.
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

```{r}

```

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

```{r}

```

